---
title: "KMeans"
author: "Melique Daley"
date: "14/04/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, include=FALSE}
x <- c("tidyquant", "PerformanceAnalytics", "timetk", "tidyr", "cluster",
       "factoextra", "ggplot2", "tidyverse", "bsts")
x <- lapply(x, require, character.only = TRUE)
```

```{r}
log_ret_monthly_xts <- read.csv("log_ret_monthly_xts.csv")
log_ret_monthly_xts <- subset(log_ret_monthly_xts, select = -c(X.1))
log_ret_monthly_xts$Date <- as.Date(log_ret_monthly_xts$Date , "%Y-%m-%d")
log_ret_monthly_xts <- tk_xts(log_ret_monthly_xts)

log_ret_monthly_sp500_xts <- read.csv("log_ret_monthly_sp500_xts.csv")
log_ret_monthly_sp500_xts <- subset(log_ret_monthly_sp500_xts, select = -c(X))
log_ret_monthly_sp500_xts$Date <- as.Date(log_ret_monthly_sp500_xts$Date , "%Y-%m-%d")
log_ret_monthly_sp500_xts <- tk_xts(log_ret_monthly_sp500_xts)
```

```{r}
profile <- read.csv("profile.csv", header=TRUE, row.names="Stocks")
profile <- profile[colnames(log_ret_monthly_xts),] 

profile$Market.Cap <- format(round(as.numeric(profile$Market.Cap), 1), nsmall=1, big.mark=",")
profile.Freq <- as.data.frame(table(subset(profile, select=c(Sector))))
profile.Country <- as.data.frame(table(subset(profile, select=c(Country))))
colnames(profile.Freq)[1] <- "Sector"
colnames(profile.Country)[1] <- "Country"
```



```{r}
ggplot(data=profile.Freq, aes(x=Sector, y=Freq)) +
  geom_bar(stat="identity", fill="steelblue")+
  geom_text(aes(label=Freq), vjust=1.6, color="white", size=3.5)+
  ggtitle("Sector Distribution")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Calculate the covariance matrix of the stocks. This is the similarity matrix . 
Annualize it by multiplying by 21 since each month has 21 trading days.
```{r}
df <- log_ret_monthly_xts
mean_ret <- colMeans(df)
cov_mat <- cov(df) * 21 
```

```{r message=FALSE}
par(mfrow=c(1,2))
fviz_nbclust(cov_mat, kmeans, method = "silhouette")
fviz_nbclust(cov_mat, kmeans, method = "wss")
```

We focus more on silhouette and pick 5 clusters

```{r}
kmeans_cluster <- kmeans(cov_mat, centers = 5, nstart = 50)
fviz_cluster(kmeans_cluster, data = cov_mat)
```

Find the stocks to be in the portfolio
```{r message=FALSE, warning=FALSE}

get_members <- function(clusters, x){
  names(clusters$cluster[clusters$cluster == x])
}

clsuter_rank <- function(df, clusters, rf){
  cluster_nums <- seq(1, length(clusters$size))
  temp <- cluster_nums
  
  for (i in cluster_nums){
    members <- get_members(clusters, i)
    ratios <- SharpeRatio(df[, members], rf=rf)[1,]
    temp[i] <- mean(ratios)
  }
  
  temp2 <- sort(temp, decreasing = TRUE)
  rank <- sapply(temp2, match, table = temp)
  return(rank)
}

SharpeRatio_finder <- function(df, clusters, x, rf, n) {
  members <- get_members(clusters, x)
  ratios <- SharpeRatio(df[, members], rf=rf)[1,]
  names(ratios[order(ratios, decreasing=TRUE)[1:n]])
}

get_stocks <- function(n, df, clusters, rf){
  c <- length(clusters$size)
  dist <- ceiling(GeometricSequence(c,n))[-1]
  dist[1] <- dist[1]-1
  index <- floor(runif(1, min = 2, max=length(dist)))
  dist[index] = dist[index] + 1
  
  dist <- c(40, 30, 20, 10)/100
  dist <- round(n*dist)
  stocks <- vector(mode = "list", length = length(dist))
  
  ranks <- clsuter_rank(df, clusters, rf)
  
  for(i in dist){
    stocks[[i]] <- SharpeRatio_finder(df, clusters, ranks[i], rf, dist[i])

  }
  
  return(unlist(stocks))
}

port <- get_stocks(8, df, kmeans_cluster, 0.01)
```

```{r}
profile[port,]
```

The following plot shows if you invested $1 in each stock how would it perform 
```{r}
df_sp500_port <- merge(df[,port], log_ret_monthly_sp500_xts)
chart.CumReturns(df_sp500_port,wealth.index=TRUE, main="Growth of $1", legend.loc="topleft")
```
We see all stocks outperform the SP500

```{r warning=FALSE}
set.seed("314159")
generate_portfolios <- function(num_port, port, rf){

  all_wts <- matrix(nrow = num_port, ncol = length(port))
  port_returns <- vector('numeric', length = num_port)
  port_risk <- vector('numeric', length = num_port)
  sharpe_ratio <- vector('numeric', length = num_port)
  
  x <- seq_along(port_returns)
  for (i in x) {
    
    wts <- runif(length(port))
    wts <- wts/sum(wts)
    
    # Storing weight in the matrix
    all_wts[i,] <- wts
    
    # Portfolio returns
    
    port_ret <- sum(wts * mean_ret[port])
    port_ret <- ((port_ret + 1)^21) - 1
    
    # Storing Portfolio Returns values
    port_returns[i] <- port_ret
    
    
    # Creating and storing portfolio risk
    port_sd <- sqrt(t(wts) %*% (cov_mat[port,port]  %*% wts))
    port_risk[i] <- port_sd
    
    # Creating and storing Portfolio Sharpe Ratios
    
    sr <- (port_ret-rf)/port_sd
    sharpe_ratio[i] <- sr
  }
  portfolio_values <- tibble(Return = port_returns, Risk = port_risk, SharpeRatio = sharpe_ratio)
  all_wts <- tk_tbl(all_wts)
  colnames(all_wts) <- port
  portfolio_values <- tk_tbl(cbind(all_wts, portfolio_values))
  return(portfolio_values)
}

portfolios <- generate_portfolios(5000, port, 0.01)
head(portfolios)
```

The minimum variance is given by
```{r}
min_var <- portfolios[which.min(portfolios$Risk),]
min_var
```

The highest sharpe ratio is given by
```{r}
max_sr <- portfolios[which.max(portfolios$SharpeRatio),]
max_sr
```

The following graph shows the weight distribution  for port
```{r}
p <- min_var %>%
  gather(all_of(port), key = Asset,
         value = Weights) %>%
  mutate(Asset = as.factor(Asset)) %>%
  ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  labs(x = 'Assets', y = 'Weights', title = "Minimum Variance Portfolio Weights") +
  scale_y_continuous(labels = scales::percent) 
p
```

The Tangency Portfolio is the optimal portfolio of risky assets
```{r}
p <- max_sr %>%
  gather(all_of(port), key = Asset,
         value = Weights) %>%
  mutate(Asset = as.factor(Asset)) %>%
  ggplot(aes(x = fct_reorder(Asset,Weights), y = Weights, fill = Asset)) +
  geom_bar(stat = 'identity') +
  theme_minimal() +
  labs(x = 'Assets', y = 'Weights', title = "Tangency Portfolio Weights") +
  scale_y_continuous(labels = scales::percent) 
p
```

The following plot shows all Portfolios that your generated 
```{r}
p <- portfolios %>%
  ggplot(aes(x = Risk, y = Return, color = SharpeRatio)) +
  geom_point() +
  theme_classic() +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent) +
  labs(x = 'Annualized Risk',
       y = 'Annualized Returns',
       title = "Generated Portfolios") +
  geom_point(aes(x = Risk,
                 y = Return), data = min_var, color = 'green') +
  geom_point(aes(x = Risk,
                 y = Return), data = max_sr, color = 'red')
p
```
Green is the minimum variance portfolio and red is tangency portfolio


```{r}
df_sp500_var_min_port <- df_sp500_port
df_sp500_var_min_port$COST <- df_sp500_var_min_port$COST * as.numeric(min_var[1])
df_sp500_var_min_port$ROL <- df_sp500_var_min_port$ROL * as.numeric(min_var[2])
df_sp500_var_min_port$WCN <- df_sp500_var_min_port$WCN * as.numeric(min_var[3])
df_sp500_var_min_port$SNPS <- df_sp500_var_min_port$SNPS * as.numeric(min_var[4])
df_sp500_var_min_port$CDNS <- df_sp500_var_min_port$CDNS * as.numeric(min_var[5])
df_sp500_var_min_port$CGNX <- df_sp500_var_min_port$CGNX * as.numeric(min_var[6])
df_sp500_var_min_port$LRCX <- df_sp500_var_min_port$LRCX * as.numeric(min_var[7])

df_sp500_tang_port <- df_sp500_port
df_sp500_tang_port$COST <- df_sp500_tang_port$COST * as.numeric(max_sr[1])
df_sp500_tang_port$ROL <- df_sp500_tang_port$ROL * as.numeric(max_sr[2])
df_sp500_tang_port$WCN <- df_sp500_tang_port$WCN * as.numeric(max_sr[3])
df_sp500_tang_port$SNPS <- df_sp500_tang_port$SNPS * as.numeric(max_sr[4])
df_sp500_tang_port$CDNS <- df_sp500_tang_port$CDNS * as.numeric(max_sr[5])
df_sp500_tang_port$CGNX <- df_sp500_tang_port$CGNX * as.numeric(max_sr[6])
df_sp500_tang_port$LRCX <- df_sp500_tang_port$LRCX * as.numeric(max_sr[7])

kmeans_port <- merge(df_sp500_tang_port[,c("GSPC", "Portfolio")], df_sp500_var_min_port[,c("Portfolio")])
colnames(kmeans_port) <- c("GSPC", "Tangency Portfolio", "Minimum Variance Portfolio")
chart.CumReturns(kmeans_port,wealth.index=TRUE, main="Growth of $1", legend.loc="topleft")
```
